---
title: "PIK GHG data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load packages
library(data.table)
library(ggplot2)
library(dplyr)
library(readxl)
library(ggrepel)
library(plotly)
library(knitr)

rm(list = ls())

# Set working directory
#setwd("C:/Users/wb460271/OneDrive - WBG/Documents/WDI-GHG")
setwd("C:/Users/wb460271/OneDrive - WBG/Documents/GitHub/WDI_GHG_emissions")

####### Read and reshape data ####### 
## PIK, downloaded from https://zenodo.org/record/5494497#.YZJPvmC0uUk # v.2.3.1 (1750-2019)
# Gütschow, J.; Günther, A.; Pflüger, M. (2021): The PRIMAP-hist national historical emissions time series v2.3.1 (1850-2019). zenodo. doi:`10.5281/zenodo.5494497`.
PIK <- as.data.table(read.csv("./Data_private/PIK/Guetschow-et-al-2021-PRIMAP-hist_v2.3.1_20-Sep_2021.csv"))
#PIK <- as.data.table(read.csv("./Data_private/PIK/PRIMAP-hist_v1.1_06-Mar-2017.csv"))

```

## PIK data description
New version of PRIMAP-hist GHG emissions data was published in September 2021 spanning data until 2019.

Gütschow, J.; Günther, A.; Pflüger, M. (2021): The PRIMAP-hist national historical emissions time series v2.3.1 (1850-2019). zenodo. doi:`10.5281/zenodo.5494497`.

Data dowloaded from https://zenodo.org/record/5494497#.YZJPvmC0uUk

### Data composition
Several datasets available:

- Guetschow-et-al-2021-PRIMAP-hist_v2.3.1_20_Sep_2021.X: The main dataset with numerical extrapolation of all time series to 2019 and three significant digits.
- Guetschow-et-al-2021-PRIMAP-hist_v2.3.1_no_extrap_20_Sep_2021.X: Variant without numerical extrapolation of missing values and not including the country groups mentioned in section [“country”] (three significant digits).
- Guetschow-et-al-2021-PRIMAP-hist_v2.3.1_no_rounding_20_Sep_2021.X: The main dataset with numerical extrapolation of all time series to 2019 and eleven significant digits.
- Guetschow-et-al-2021-PRIMAP-hist_v2.3.1_no_extrap_no_rounding_20_Sep_2021.csv: Variant without numerical extrapolation of missing values and not including the country groups mentioned in section [“country”] (eleven significant digits).

For this exercise, the first dataset with numerical extrapolation is used. 

The dataset includes data for two different scenarios:

- HISTCR: In this scenario country-reported data (CRF, BUR, UNFCCC) is prioritized over third-party data (CDIAC, FAO, Andrew, EDGAR, BP).
- HISTTP: In this scenario third-party data (CDIAC, FAO, Andrew, EDGAR, BP) is prioritized over country-reported data (CRF, BUR, UNFCCC))

Both series are considered.

### Temporal coverage
The dataset covers the years 1750-2019, where a considerable number of values 2019 are numerically extrapolated. As the WDI starts in 1960, only values from 1960 are considered.

### Country coverage
```{r country_coverage, include=FALSE}
# Check which WDI countries available
# Load country WDI list (217 countries)
WDI_countries <- as.data.table(read.csv("Data/wdi_country_list.csv"))
setnames(WDI_countries, c("long_name", "ISO3"))

PIK_countries <- unique(PIK$area..ISO3.)
length(PIK_countries) # 215 countries
table(WDI_countries$ISO3 %in% PIK_countries) # 198 WDI countries available in PIK
WDI_countries[!(ISO3 %in% PIK_countries)] # list of missing countries
table(PIK_countries %in% WDI_countries$ISO3)
PIK_countries[!(PIK_countries %in% WDI_countries$ISO3)] # list of missing countries

# Drop countries/regions not in WDI
PIK <- PIK[area..ISO3. %in% WDI_countries$ISO3]
```

The WDI includes `r nrow(WDI_countries)` of which `r sum(WDI_countries$ISO3 %in% PIK_countries)` are included in the PIK dataset. The following WDI countries are not included in the PIK dataset:

```{r echo = FALSE, results='asis'}
kable(WDI_countries[!(ISO3 %in% PIK_countries)])
```

# View country-by-country data
```{r dataprep, echo = FALSE, results='asis'}
# Convert data in long format
PIK_long <- melt(PIK, id.vars = c("source", "ISO3", "entity", "unit", "category", "scenario"),
                 variable.name = "year", value.factor = FALSE, variable.factor = FALSE)
PIK_long[, year := as.numeric(substring(year, 2, 5))]
dim(PIK_long)
```

## CO2



