---
title: "PIK GHG data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Load packages
library(data.table)
library(ggplot2)
library(dplyr)
library(readxl)
library(ggrepel)
library(plotly)
library(knitr)
library(htmltools)

rm(list = ls())

# Set working directory
#setwd("C:/Users/wb460271/OneDrive - WBG/Documents/WDI-GHG")
setwd("C:/Users/wb460271/OneDrive - WBG/Documents/GitHub/WDI_GHG_emissions")

####### Read data ####### 
## PIK, downloaded from https://zenodo.org/record/5494497#.YZJPvmC0uUk # v.2.3.1 (1750-2019)
# Gütschow, J.; Günther, A.; Pflüger, M. (2021): The PRIMAP-hist national historical emissions time series v2.3.1 (1850-2019). zenodo. doi:`10.5281/zenodo.5494497`.
PIK <- as.data.table(read.csv("./Data_private/PIK/Guetschow-et-al-2021-PRIMAP-hist_v2.3.1_20-Sep_2021.csv"))
#PIK <- as.data.table(read.csv("./Data_private/PIK/PRIMAP-hist_v1.1_06-Mar-2017.csv"))
setnames(PIK, "scenario..PRIMAP.hist.", "scenario")

```

## PIK data description
New version of PRIMAP-hist GHG emissions data was published in September 2021 spanning data until 2019.

Gütschow, J.; Günther, A.; Pflüger, M. (2021): The PRIMAP-hist national historical emissions time series v2.3.1 (1850-2019). zenodo. doi:`10.5281/zenodo.5494497`.

Data dowloaded from https://zenodo.org/record/5494497#.YZJPvmC0uUk

### Data composition
Several datasets available:

- Guetschow-et-al-2021-PRIMAP-hist_v2.3.1_20_Sep_2021.X: The main dataset with numerical extrapolation of all time series to 2019 and three significant digits.
- Guetschow-et-al-2021-PRIMAP-hist_v2.3.1_no_extrap_20_Sep_2021.X: Variant without numerical extrapolation of missing values and not including the country groups mentioned in section [“country”] (three significant digits).
- Guetschow-et-al-2021-PRIMAP-hist_v2.3.1_no_rounding_20_Sep_2021.X: The main dataset with numerical extrapolation of all time series to 2019 and eleven significant digits.
- Guetschow-et-al-2021-PRIMAP-hist_v2.3.1_no_extrap_no_rounding_20_Sep_2021.csv: Variant without numerical extrapolation of missing values and not including the country groups mentioned in section [“country”] (eleven significant digits).

For this exercise, the first dataset with numerical extrapolation is used. 

The dataset includes data for two different scenarios:

- HISTCR: In this scenario country-reported data (CRF, BUR, UNFCCC) is prioritized over third-party data (CDIAC, FAO, Andrew, EDGAR, BP).
- HISTTP: In this scenario third-party data (CDIAC, FAO, Andrew, EDGAR, BP) is prioritized over country-reported data (CRF, BUR, UNFCCC))

Both series are considered.

### Temporal coverage
The dataset covers the years 1750-2019, where a considerable number of values 2019 are numerically extrapolated. As the WDI starts in 1960, only values from 1960 are considered.

### Country coverage
```{r country_coverage, include=FALSE}
# Check which WDI countries available
# Load country WDI list (217 countries)
WDI_countries <- as.data.table(read.csv("Data/wdi_country_list.csv"))
setnames(WDI_countries, c("long_name", "ISO3"))

PIK_countries <- unique(PIK$area..ISO3.)
length(PIK_countries) # 215 countries
table(WDI_countries$ISO3 %in% PIK_countries) # 198 WDI countries available in PIK
WDI_countries[!(ISO3 %in% PIK_countries)] # list of missing countries
table(PIK_countries %in% WDI_countries$ISO3)
PIK_countries[!(PIK_countries %in% WDI_countries$ISO3)] # list of missing countries

# Drop countries/regions not in WDI
PIK <- PIK[area..ISO3. %in% WDI_countries$ISO3]

# Rename variables
colnames(PIK)
setnames(PIK, c("area..ISO3.", "category..IPCC2006_PRIMAP."),
         c("ISO3", "category"))
```

The WDI includes `r nrow(WDI_countries)` of which `r sum(WDI_countries$ISO3 %in% PIK_countries)` are included in the PIK dataset. The following WDI countries are not included in the PIK dataset:

```{r echo = FALSE, results='asis'}
kable(WDI_countries[!(ISO3 %in% PIK_countries)])
```

# View country-by-country data
```{r dataprep, echo = FALSE, results='asis'}
# Convert data in long format
PIK_long <- melt(PIK, id.vars = c("source", "ISO3", "entity", "unit", "category", "scenario"),
                 variable.name = "year", value.factor = FALSE, variable.factor = FALSE)
PIK_long[, year := as.numeric(substring(year, 2, 5))]
#dim(PIK_long)
PIK_long <- PIK_long[year >= 1960, ] # only keep if year >= 1960
PIK_long[, value := value / 1000] # convert to Mt CO2e

## CAIT data for comparison
CAIT <- as.data.table(read_xlsx("Data_private/CAIT/ghg-emissions/CW_CAIT_GHG_Emissions.xlsx"))
#CAIT

CAIT_long <- melt(CAIT, id.vars = c("Country", "Source", "Sector", "Gas"))
setnames(CAIT_long, "value", "CO2_emissions_ktCO2")
setnames(CAIT_long, "variable", "Year")
CAIT_long[, Source := NULL]
CAIT_long[, Year := as.numeric(as.character(Year))] # convert year to numeric

CAIT_wide <- dcast(data = CAIT_long %>% filter(Sector == "Total excluding LUCF" & Gas == "CO2"), #  & Gas == "All GHG"
                   formula = Year ~ Country,
                   value.var = "CO2_emissions_ktCO2")
#class(CAIT_wide$Year)
CAIT_wide[, Year := as.numeric(as.character((Year)))]

all_ISO3 <- intersect(unique(PIK_long$ISO3), unique(CAIT_long$Country)) # c("RUS", "NLD")

##### Plot series for each country #####
plots <- lapply(all_ISO3, function(cur_ISO3){
  #cur_ISO3 <- "RUS"
  
  # Select only current country
  cur_PIK_long <- PIK_long %>% filter(ISO3 == cur_ISO3)
  
  # Reshape to wide
  cur_PIK_wide <- cur_PIK_long %>% select(-unit) %>% 
    dcast(formula = ... ~ entity + scenario + category) 
  #cur_PIK_wide
  
  # Add CAIT data
  cur_CAIT_wide <- CAIT_long %>% 
    mutate(Sector = gsub(" ", "", Sector)) %>% # remove white spaces in sector names
    filter(Country == cur_ISO3) %>% 
    dcast(formula = ... ~ Gas + Sector, value.var = "CO2_emissions_ktCO2") %>%
    rename(year = Year, ISO3 = Country)
  
  #sapply(cur_CAIT_wide, class)
  #sapply(cur_PIK_wide, class)
  
  cur_combined_wide <- merge(cur_CAIT_wide, cur_PIK_wide, by = c("ISO3", "year"), all.y = TRUE)
  dim(cur_combined_wide)
  
  cur_fig <- plot_ly(cur_combined_wide, x = ~year, y = ~CO2_HISTCR_M.0.EL, 
                     name = 'PIK - HISTCR M.0.EL', type = 'scatter', mode = 'lines') %>% 
    add_trace(y = ~CO2_HISTTP_M.0.EL, name = 'PIK - HISTTP M.0.EL', mode = 'lines') %>%
    add_trace(y = ~CO2_TotalexcludingLUCF, name = 'CAIT - CO2_Total excluding LUCF', mode = 'lines') %>%
  layout(title = cur_ISO3, yaxis = list(title = list(text ='CO2 emissions (Mt CO2)')))
  
  #print(cur_fig)
})
htmltools::tagList(plots)
```

## CO2



